---
output:
  html_document: default
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Peer Grader Guidance
Please review the student expectations for peer review grading and peer review comments.  Overall, we ask that you score with accuracy. When grading your peers, you will not only learn how to improve your future homework submissions but you will also gain deeper understanding of the concepts in the assignments. When assigning scores, consider the responses to the questions given your understanding of the problem and using the solutions as a guide. Moreover, please give partial credit for a concerted effort, but also be thorough. **Add comments to your review, particularly when deducting points, to explain why the student missed the points.** Ensure your comments are specific to questions and the student responses in the assignment.


# Background

You have been contracted as a healthcare consulting company to understand the factors on which the pricing of health insurance depends. 

## Data Description

The data consists of a data frame with 1338 observations on the following 7 variables:

1. price: Response variable ($)
2. age: Quantitative variable
3. sex: Qualitative variable
4. bmi: Quantitative variable
5. children: Quantitative variable
6. smoker: Qualitative variable
7. region: Qualitative variable

## Instructions on reading the data

To read the data in `R`, save the file in your working directory (make sure you have changed the directory if different from the R working directory) and read the data using the `R` function `read.csv()`

```{r}
insurance = read.csv("insurance.csv", head = TRUE)
```

# Question 1: Exploratory Data Analysis [12 points]

a. **3 pts** Create plots of the response, *price*, against three quantitative predictors *age*, *bmi*, and *children*. Describe the general trend (direction and form) of each plot. 
```{r}
plot(insurance$age,insurance$price , xlab='Age',ylab='Price', col="darkblue")
plot(insurance$bmi,insurance$price , xlab='bmi',ylab='Price', col="darkblue")
plot(insurance$children,insurance$price , xlab='children',ylab='Price', col="darkblue")

```
From the price vs age scatter plot, we can find that the trend of insurance price increase when age increase. A clear linear relationship is noticed. 
From the price vs BMI scatter plot, we can determine a weak positive trending that higher BMI tends to have higher insurance price. 
Form the price vs children scatter plot, no clear pattern is notice between children 0-3, there is a drop in price when children is 4&5.




b. **3 pts** What is the value of the correlation coefficient for each of the above pair of response and predictor variables? What does it tell you about your comments in part (a).
```{r}

cor(insurance[,c(7,1,3,4)])
```
The correlation between price and age is 0.2990, between price and bmi is 0.1983, between price and children is 0.067.
The largest correlation coefficient is 0.2990 between price and age which agree with our findings in Part(a)
Overall, no strong correlation among the predicting variables are noticed.

c. **3 pts** Create box plots of the response, *price*, and the three qualitative predictors *sex*, *smoker*, and *region*. Based on these box plots, does there appear to be a relationship between these qualitative predictors and the response?

```{r}
boxplot(price~sex,main="",xlab="Sex",ylab="=Insurance Price",col=blues9,data=insurance)
boxplot(price~smoker,main="",xlab="Smoker",ylab="=Insurance Price",col=blues9,data=insurance)
boxplot(price~region,main="",xlab="Region",ylab="=Insurance Price",col=blues9,data=insurance)
```
From plot 1, sex does not really affect the insurance price. 
Form plot 2, non-smokers have much lower insurance price. 
From plot 3, region is not a strong factor to insurance price.

*Hint*: Use the given code to convert the qualitative predictors to factors.

```{r}
#make categorical variables into factors
insurance$sex<-as.factor(insurance$sex) #makes female the baseline level
insurance$smoker<-as.factor(insurance$smoker) #makes no the baseline level
insurance$region<-as.factor(insurance$region) #makes northeast the baseline level
```


d. **3 pts** Based on the analysis above, does it make sense to run a multiple linear regression with all of the predictors?

I think it make sense to run with all of the predictors. From part a&b, we can conclude all predictors contribute to the insurance price in some level. No evidence shows that we should exclude any of predictor here. 


*Note: Please work on non-transformed data for all of the following questions.*

# Question 2: Fitting the Multiple Linear Regression Model [10 points]

Build a multiple linear regression model, named *model1*, using the response, *price*, and all 6 predictors, and then answer the questions that follow:

a. **5 pts** Report the coefficient of determination for the model and give a concise interpretation of this value.
```{r}
model1 = lm(price ~ .,data = insurance)
summary(model1)

```
From the model summary, there are 9 predictor rows in addition to the intercept.Smokeryes has the largest coefficient which significant increase the insurance price. Southeast region reduce the predict insurance price. To reivew the p-value, age, bmi, children and smokeryes are four most significant predictor.



b. **5 pts** Is the model of any use in predicting price? Conduct a test of overall adequacy of the model, using $\alpha = 0.05$. Provide the following elements of the test: null hypothesis $H_0$, alternative hypothesis $H_a$, F- statistic or p-value, and conclusion.

H0: The coefficients corresponding to the predictors are all zero. 
HA: At least on coefficient is not zero. 

For overall regression, the F-value is 500.8 and the P-value is approximately zero. He reject H0 which means that at least one of the predictive variables has predictive power.



# Question 3: Model Comparison [12 points]

a. **4 pts** Assuming a marginal relationship between *region* and *price*, perform an ANOVA F-test on the mean insurance prices among the different regions. Using an $\alpha-level$ of 0.05, can we reject the null hypothesis that the means of the regions are equal? Please interpret.
```{r}
reducemodel = aov(price ~ region, data = insurance)
summary(reducemodel)
```
From the above results, it is observed that the F-statistic value is 2.97 and it is significant as the corresponding p-value 0.0309 which is less than the level of significance (0.05). Thus, it is wise to reject the null hypothesis of equal mean value of all regions. Different regions has different coefficients. 



b. **4 pts** Now, build a second multiple linear regression model, called *model2*, using *price* as the response variable, and all variables except *region* as the predictors. Conduct a partial F-test comparing *model2* with *model1*. What is the partial-F test p-value? Can we reject the null hypothesis that the regression coefficients for *region* variables are zero at $\alpha-level$ of 0.05?
```{r}
model2 = lm(price ~ age + sex + bmi + children +smoker,data = insurance)
summary(model2)
anova(model2,model1)
```
Partial F test:
F-value: 2.1173, P-value: 0.09622

Because the P-value is larger than 0.05, we cannot reject the null hypothesis that the regression coefficients for regions are zero given all other predictors in model1, atα-level of 0.05.




c. **4 pts** What can you conclude from a and b? Do they provide the exact same results? 

From part a, we conclude different regions contribute different to the insurance price. 
From part b, we conclude adding Region predictor does not contribute to the change in insurance price. 
The above two analysis give opposite answers. 




*Note: Please use model1 for all of the following questions.*

# Question 4: Coefficient Interpretation [6 points]

a. **3 pts** Interpret the estimated coefficient of *sexmale* in the context of the problem. *Make sure female is the baseline level for sex. Mention any assumption you make about other predictors clearly when stating the interpretation.*

The estimated coefficient for sexmale is -131.3.
Based on this model, it seems to say that compare to female (as the baseline control), male trends to have a lower insurance price. If we keep other predictors as constant, This conclusion can be true because woman may need more frequent medical care. 


b. **3 pts** If the value of the *bmi* in *model1* is increased by 0.01 keeping other predictors constant, what change in the response would be expected?
```{r}
summary(model1)
```
The estimated coefficient of BMI is 339.2. If we keep other predictors constant, a increasement of 0.01 in BMI will cause the insurance price increase 3.39. 


# Question 5: Confidence and Prediction Intervals [10 points]

a. **5 pts** Compute 90% and 95% confidence intervals (CIs) for the parameter associated with *age* for *model1*. What observations can you make about the width of these intervals?
```{r}
confint(model1, "age", level = 0.95)
```
This bounds do not make sense as human are not likely to live over 200 yrs old. 


b. **2.5 pts** Using *model1*, estimate the average price for all insurance policies with the same characteristics as the first data point in the sample. What is the 95% confidence interval? Provide an interpretation of your results.
```{r}
newdata = insurance[1,]
predict(model1, newdata, interval="confidence") 
```
The fit value is 25293.71. The 95% confidence interval is (24143.98,26443.44). For policies with the same characteristics as the first data point, the average estimated price are $25293.71, with a lower bound of $24143.98 and an upper bound of $26443.44. 

c. **2.5 pts** Suppose that the *age* value for the first data point is increased to 50, while all other values are kept fixed. Using *model1*, predict the price of an insurance policy with these characteristics. What is the 95% prediction interval? Provide an interpretation of your results.
```{r}
xstar.new = insurance
xstar.new[1,1] = 50
predict(model1, xstar.new[1,], interval="prediction")
```
The fit value is 33256.26. The 95% confidence interval is (21313.29, 45199.23). For policies with the same characteristics as this data point, the average estimated price are $33256.26., with a lower bound of $21313.29 and an upper bound of $45199.23 

